---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# The Ranking Project: Visualizations for Comparing Populations

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/civilstat/RankingProject.svg?branch=master)](https://travis-ci.org/civilstat/RankingProject)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/RankingProject)](https://cran.r-project.org/package=RankingProject)
<!-- badges: end -->

The package `RankingProject` is a companion for the articles ["A Primer on Visualizations for Comparing Populations, Including the Issue of Overlapping Confidence Intervals"](https://doi.org/10.1080/00031305.2017.1392359) (Wright, Klein, and Wieczorek, 2019, *The American Statistician*) and ["A Joint Confidence Region for an Overall Ranking of Populations"](https://doi.org/10.1111/rssc.12402) (Klein, Wright, and Wieczorek, 2020, *Journal of the Royal Statistical Society: Series C*).

The package provides functions for plotting ranked tables of data side-by-side with their plots. The available visualizations include shaded columns plots, adjusted confidence intervals, and related plots intended for making correct inferences about one-to-many or many-to-many comparisons.

## Installation

```{r, eval = FALSE}
# install.packages("devtools")
# library("devtools")
install_github("civilstat/RankingProject")
```

To install the vignettes and exactly replicate figures from the paper,
you will also need the `tikzDevice` package:

```{r, eval = FALSE}
# install.packages("tikzDevice")
install_github("civilstat/RankingProject", build_vignettes = TRUE)
```

## Example

```{r, fig.height=8, fig.width=6.5}
library(RankingProject)

# Load dataset of mean travel time (in minutes) to work
# of workers 16 years and over who did not work at home,
# from the 2011 American Community Survey (ACS)
data(TravelTime2011)
USdata <- TravelTime2011
head(USdata)

# Format estimates and SEs into strings with 2 digits past the decimal
USdata$Estimate.Print = formatC(USdata$Estimate.2dec,
                                format = 'f', digits = 2)
# For SEs, also drop the leading 0
USdata$SE.Print = substring(formatC(USdata$SE.2dec,
                                    format = 'f', digits = 2),
                            first = 2)

# Set Colorado as the reference area
refAbbr <- "CO"
refRow  <- which(USdata$Abbreviation==refAbbr)

# Set up parameter lists for table function and plot function
tableParList <- with(USdata,
                     list(ranks = Rank, names = State,
                          est = Estimate.Print, se = SE.Print,
                          placeType = "State"))
plotParList <- with(USdata,
                      list(est = Estimate.2dec, se = SE.2dec,
                           names = Abbreviation, refName = refAbbr,
                           confLevel = .9, cex = 0.6, thetaLine = 1.5,
                           plotType = "difference", lwdBold = 2))

# Plot demi-Bonferroni-corrected 90% CIs
# for the differences between Colorado and each other state,
# alongside a ranking table
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList)
```

## References

Almond, R.G., Lewis, C., Tukey, J.W., and Yan, D. (2000). "Displays for Comparing a Given State to Many Others," *The American Statistician*, vol. 54, no. 2, 89-93, DOI:10.1080/00031305.2000.10474517.

Klein, M., Wright, T., and Wieczorek, J. (2020). "A Joint Confidence Region for an Overall Ranking of Populations," *Journal of the Royal Statistical Society: Series C*, vol. 69, no. 3, 589-606, DOI:10.1111/rssc.12402.

Wright, T., Klein, M., and Wieczorek, J. (2019). "A Primer on Visualizations for Comparing Populations, Including the Issue of Overlapping Confidence Intervals," *The American Statistician*, vol. 73, no. 2, 165-178, DOI:10.1080/00031305.2017.1392359.

