---
title: "Figures for 'A Primer on Visualizations for Comparing Populations...'"
author: "Jerzy Wieczorek"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Figures for 'A Primer on Visualizations for Comparing Populations...'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
  - \usepackage{tikz}
  - \usepackage{preview}
---

```{r, echo=FALSE, cache=FALSE}
library(knitr)
options(tikzMetricsDictionary="tikzDictionary",
        tikzMetricPackages = c("\\usepackage[utf8]{inputenc}","\\usepackage[T1]{fontenc}",
                               "\\usetikzlibrary{calc}", "\\usepackage{amssymb}"))
opts_chunk$set(fig.width=6.5, fig.height=8, cache=FALSE, message=FALSE,
               dev='tikz', external=FALSE,
               out.width=6.5, out.height=8)
# Fixed issue with dev='tikz' in Windows
# by setting external=FALSE and a new plot hook, as here:
# https://github.com/yihui/tikzDevice/issues/60
knit_hooks$set(crop = NULL, plot = function(x, options) {
  if ('tikz' %in% options$dev && !options$external) {
    hook_plot_tex(x, options)
  } else hook_plot_md(x, options)
})
```

This vignette shows how to reproduce the main figures in ["A Primer on Visualizations for Comparing Populations, Including the Issue of Overlapping Confidence Intervals"](https://doi.org/10.1080/00031305.2017.1392359) (Wright, Klein, and Wieczorek, 2019, *The American Statistician*).

Note: For this vignette itself, we automatically save the figures below using the `knitr` package with option `dev="tikz"` instead of saving them individually. In the final section of the vignette, we show an example of how to save individual plots using the `tikz()` function in the `tikzDevice` package.

## Workflow to reproduce figures from the article

First, we load the package and the `TravelTime2011` dataset used in the paper. We also create string versions of our estimates and their standard errors that will print with a consistent number of digits.

```{r}
library(RankingProject)
data(TravelTime2011)
USdata <- TravelTime2011
head(USdata)

# Format estimates and SEs into strings with 2 digits past the decimal
USdata$Estimate.Print = formatC(USdata$Estimate.2dec,
                                format = 'f', digits = 2)
# For SEs, also drop the leading 0
USdata$SE.Print = substring(formatC(USdata$SE.2dec,
                                    format = 'f', digits = 2),
                            first = 2)
```

Next, we set up several list-type objects to contain parameters needed for the tables and plots. As in the article, we use Colorado (CO) as the reference state. The option `tikzText=TRUE` lets us use LaTeX-style text and symbols in the figures, instead of basic R-style text.

```{r}
# Set Colorado as the reference state
refAbbr <- "CO"
refRow  <- which(USdata$Abbreviation==refAbbr)

# Set up parameter lists for table function and figure function
tableParList <- with(USdata,
                     list(ranks = Rank, names = Abbreviation, 
                          est = Estimate.Print, se = SE.Print,
                          placeType = "State", tikzText = TRUE))
plotParList <- with(USdata,
                      list(est = Estimate.2dec, se = SE.2dec,
                           names = Abbreviation, refName = refAbbr,
                           confLevel = .90, tikzText = TRUE))
```

Reproduce Figure 3, the "shaded columns plot":

```{r}
# Shaded Columns plot
plotParList$plotType <- "columns"
# Specify where to position the "Reference State:" text,
# and adjust column widths from their defaults
tableParList = c(tableParList,
                 list(columnsPlotRefLine = .7, col2 = .55, col3 = .8))
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tableWidthProp = 2/7, tikzText = TRUE)
```

```{r}
# Reset defaults for future plots
tableParList[c("columnsPlotRefLine", "col2", "col3")] <- NULL
```

Reproduce Figure 1, the plot of individual 90% confidence intervals (CIs):

```{r}
# For all remaining figures,
# table will show full state names instead of abbreviations
tableParList$names <- USdata$State

# Individual CIs
plotParList$plotType <- "individual"
plotParList$cex <- 0.6
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
```

Reproduce Figure 4, the plot of demi-Bonferroni-corrected 90% CIs for the difference between the reference state Colorado and all other states:

```{r}
# CIs for differences from ref
plotParList$plotType <- "difference"
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  annotRefName = USdata$State[refRow],
                  annotRefRank = USdata$Rank[refRow],
                  tikzText = TRUE)
```

Reproduce Figure 7, the plot of demi-Bonferroni-corrected 90% "comparison intervals" (based on Almond et al., 2000) for comparing the reference state Colorado to all other states:

```{r}
# Comparison intervals
plotParList$plotType <- "comparison"
plotParList$thetaLine <- 1.5
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
plotParList$thetaLine <- NULL
```

Reproduce Figure 10, the plot of Goldstein-Healy-adjusted "90%" CIs (based on Goldstein and Healy, 1995), which are in fact 77.49% CIs as chosen to achieve an "average significance level" of $\alpha=0.10$:

```{r}
# Goldstein-Healy adjusted CIs
plotParList$plotType <- "individual"
plotParList$GH <- TRUE
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
```

Reproduce Figure 11, the plot of two-tiered error bars, where the inner tier are the same Goldstein-Healy-adjusted "90%" CIs from Figure 10, and the outer tier are the same individual 90% CIs from Figure 1:

```{r}
# Double-tiered GH plot:
# inner tiers are GH CIs,
# outer tiers are usual 90% CIs
plotParList$tiers <- 2
# Legend auto-positioning is poor with line breaks in legend text;
# we can improve it by controlling (X,Y) manually
plotParList$legendX <- 13
plotParList$legendY <- 52
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
```

Reproduce Figure 12, another plot of two-tiered error bars, where now the inner tier are the same individual 90% CIs from Figure 1, and the outer tier are demi-Bonferroni-corrected Goldstein-Healy-adjusted "90%" CIs (in fact 99.55% CIs):

```{r}
# Double-tiered GH + Bonferroni plot:
# inner tiers are usual 90% CIs,
# outer tiers are 50-way demi-Bonferroni-corrected GH CIs
plotParList$Bonferroni <- "demi"
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
```

## Using `tikzDevice::tikz()` to save individual figures

For this vignette, the figures above were automatically converted to PDF format using `knitr` with chunk option `dev="tikz"`. When not using `knitr`, we may prefer to save plots one at a time "manually." To do this, we can explicitly call the `tikz()` function from the `tikzDevice` package, as in the following example code.

The `tikz()` function works much like `pdf()` or `png()` and other standard functions for saving plots from R scripts. We must remember to call `dev.off()` after the plotting function runs, to let R know the plot is ready to be saved.

Using `tikz()` will create and save a `.tex` file. To convert this to a figure, we can:

* compile it into a standalone PDF separately; or
* use R's `tools::texi2pdf()` which compiles the PDF and saves it in the current working directory; or
* set `standAlone=FALSE` below, then copy-paste the contents of the saved `.tex` file directly into a larger `.tex` document.

```{r, eval=FALSE}
# Not run:
library(tikzDevice)
tikz("/path/to/my/file.tex", standAlone = TRUE, width = 6.5, height = 8)
RankPlotWithTable(tableParList = tableParList, plotParList = plotParList,
                  tikzText = TRUE)
dev.off()
tools::texi2pdf("/path/to/my/file.tex")
```

